---
title: "Vignette for tPSA Example"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
```

In this file we reproduce the results of the tPSA example presented in Section 3 of the paper. This example concerns the comparison of methods of measuring total Prostatic Specific Antigen (tPSA,
μg/L) measurements as first reported in [Ferraro et al. (2023)](https://www.degruyterbrill.com/document/doi/10.1515/cclm-2022-0874/html). The various analyses presented here make use of the functions provided in the file `functions.R` and the data provided in the file `tPSADataRvsS.csv`. We begin by loading those functions and loading that data.

### Load the data and necessary functions

```{r}
source(file = "functions.R")
svr <- read.csv(file = "tPSADataRvsS.csv", header = TRUE)
X <- svr$MS1
r_vec_x <- rep(2, 135)
Y <- svr$MS2
r_vec_y <- rep(2, 135)
```

### Visualize the data

Next we visualize the data through various scatterplots. The call below to `data_visualization()` reproduces the plots provided in Figure 1 in the paper. 

```{r, fig.width=8, fig.height=8, fig.align='center'}
data_vizualization(x = X, r_x = r_vec_x, y = Y, r_y = r_vec_y, s_name = "tPSA", ref_name = "Roche", com_name = "Siemens")
```

### Model fitting

We next fit the model described by Equation (2) in the paper. Note that we do not specify the desired polynomial orders; instead we perform order selection via the BIC-based selection method described in Section 2.1 of the paper. This is achieved with the `fit_model()` function and the `order_select` input set to `TRUE`. Having done that, the values passed into `orders` are taken to be the *largest* orders considered in the BIC-based selection process. Alternatively, when `order_select = FALSE`, the values passed into `orders` are the polynomial orders $(p,d_x,d_y)$ used to fit the bias and precision polynomials.

```{r, cache=TRUE}
fitted_model <- fit_model(x = X, r_x = r_vec_x, y = Y, r_y = r_vec_y, orders = c(4,4,4), order_select = TRUE)
print(fitted_model)
```

### Extract parameter estimates and calculate CIs

Having fit the model and obtained parameter estimates, we next calculate 95% confidence intervals for the model parameters. We do so using the *standard* bootstrap approach described in Section 2.4.1. This is achieved by first using the `do_bootstrap()` function to take the bootstrap samples, and then with the `calc_param_CIs()` function with the `type` input set to `"standard"` to calculate the necessary confidence limits. Alternatively, one could set `type = "percentile"` to calculate the percentile-based bootstrap CIs described in Section 2.4.2. The resulting table below is a reproduction of Table 1 from the paper.

```{r, cache=TRUE}
library(doParallel, quietly = TRUE)
numCores <- detectCores()
registerDoParallel(numCores)
set.seed(123456789)
res_boot <- do_bootstrap(B = 10000, x = X, r_x = r_vec_x, y = Y, r_y = r_vec_y, orders = c(1,1,2))
CI <- calc_param_CIs(boot = res_boot, alpha = 0.05, theta_hat = fitted_model$Estimates, type = "standard")
parameters <- c("mu", "sigma", paste0("beta", 0:1), paste0("omegax", 0:1), paste0("omegay", 0:2))
results <- data.frame(Parameter = parameters, Estimates = fitted_model$Estimates, LB = CI$LB, UB = CI$UB)
kable(results)
```

### Visualize model diagnostics

As demonstrated below, the function `model_diagnostics()` may be used to construct scatterplots, histograms, and QQ-plots which facilitate informal evaluation of the appropriateness of the model for each method. Note that the plots below were not included in the main text of the paper.

```{r, fig.width=8, fig.height=6, fig.align='center'}
library(car, quietly = TRUE)
model_diagnostics(x = X, r_x = r_vec_x, y = Y, r_y = r_vec_y, fitted_model = fitted_model, orders = c(1,1,2))
```

### Visualize method comparison

Having fit the model, we can now calculate and visualize various agreement metrics that quantify the comparability of the methods under comparison. With the `calculate_bias()` and `calculate_precision()` functions, we produce plots of the estimated bias and precision polynomials $g(s;\hat{\boldsymbol{\beta}})$, $\sigma_x(s;\hat{\boldsymbol{\omega}}_x)$, $\sigma_y(s;\hat{\boldsymbol{\omega}}_y)$ described in Section 2.1. The plots below are a reproduction of Figure 2 in the paper.

```{r, fig.width=10, fig.height=3.75, fig.align='center'}
par(mfrow=c(1,3))
bias <- calculate_bias(theta = fitted_model$Estimates, orders = c(1,1,2), 
                       s = seq(from = min(fitted_model$BLA), to = max(fitted_model$BLA), length.out = 500),
                       interval_info = list(type = "standard", boot = res_boot, alpha = 0.05), 
                       plot_info = list(leg_location = "bottomleft", colour = "red", s_name = "tPSA (μg/L)"))
precision_x <- calculate_precision(theta = fitted_model$Estimates, orders = c(1,1,2), method = "R", 
                       s = seq(from = min(fitted_model$BLA), to = max(fitted_model$BLA), length.out = 500),
                       interval_info = list(type = "standard", boot = res_boot, alpha = 0.05), 
                       plot_info = list(leg_location = "topleft", colour = "blue", s_name = "tPSA (μg/L)", 
                                        method_name = "Roche", ylimits = c(0, 1.6)))
precision_y <- calculate_precision(theta = fitted_model$Estimates, orders = c(1,1,2), method = "C", 
                       s = seq(from = min(fitted_model$BLA), to = max(fitted_model$BLA), length.out = 500),
                       interval_info = list(type = "standard", boot = res_boot, alpha = 0.05), 
                       plot_info = list(leg_location = "topleft", colour = "purple", s_name = "tPSA (μg/L)", 
                                        method_name = "Siemens", ylimits = c(0, 1.6)))
```

We next calculate and visualize the probability of agreement function using the `calculate_poa()` function. Note that the `delta` input controls the specification of the equivalence margin. When an indifference region with a fixed equivalence margin is desired, e.g., `[-x,x]`, one should input a scalar such as `delta = x`. Alternatively, if an indifference region with a relative equivalence margin is desired, e.g., 15%, one should input a string such as `delta = "x%"`. The plots below reproduce the left column of plots in Figure 2 in the paper.

```{r, fig.width=10, fig.height=5, fig.align='center'}
s <- seq(from = min(fitted_model$BLA), to = max(fitted_model$BLA), length.out = 500)
par(mfrow = c(1,2))
# Fixed delta
poa_fix <- calculate_poa(theta = fitted_model$Estimates, delta = 4, orders = c(1,1,2), s = s,
                       interval_info = list(type = "standard", boot = res_boot, alpha = 0.05), 
                       plot_info = list(leg_location = "bottomleft", colour = "orange", s_name = "tPSA (μg/L)"))
title(main = bquote("PoA Plot ("*delta*"=4)"))
# Relative delta
poa_rel <- calculate_poa(theta = fitted_model$Estimates, delta = "15%", orders = c(1,1,2), s = s,
                       interval_info = list(type = "standard", boot = res_boot, alpha = 0.05), 
                       plot_info = list(leg_location = "bottomright", colour = "orange", s_name = "tPSA (μg/L)"))
title(main = bquote("PoA Plot ("*delta*"=15%)"))
```

Note that the bias, precision, and PoA plots presented above were constructed with `"standard"` bootstrap confidence intervals/bands. One could instead set `type = "percentile"` in the respective calls to `calculate_bias()`, `calculate_precision()`, and `calculate_poa()` to construct percentile-based bootstrap confidence intervals/bands. Doing so would recreate Figure A1 from the Appendix.

As discussed in Section 2.5, one may wish to work with the *conditional* probability of agreement, which calculates the PoA for the individual subjects in the study. The `conditional_poa()` may be used for this purpose. Note that the `delta` input works the same way in this function as in `calculate_poa()`. The plots below reproduce the right column of plots in Figure 2 in the paper.

```{r, cache=TRUE, fig.width=10, fig.height=5, fig.align='center'}
par(mfrow = c(1,2))
# Fixed delta
numCores <- detectCores()
registerDoParallel(numCores)
set.seed(123456789)
cond_poa <- conditional_poa(x = X, r_x = r_vec_y, y = Y, r_y = r_vec_y, orders = c(1,1,2), 
                            delta = 4, B = 1000, alpha = 0.05,
                            plot_info = list(colour = "orange", s_name = "Estimated tPSA (μg/L)"))
title(main = bquote("Conditional PoA Plot ("*delta*"=4)"))
# Relative delta
cond_poa <- conditional_poa(x = X, r_x = r_vec_y, y = Y, r_y = r_vec_y, orders = c(1,1,2), 
                            delta = "15%", B = 1000, alpha = 0.05,
                            plot_info = list(colour = "orange", s_name = "Estimated tPSA (μg/L)"))
title(main = bquote("Conditional PoA Plot ("*delta*"=15%)"))
```

### Calibrated analysis

As presented in Section 3, one may wish to calibrate (bias correct) the measurements by the comparator method and then quantify agreement *after* calibration. We illustrate such an analysis here, beginning by performing the calibration.

```{r, cache=TRUE}
beta_hat <- fitted_model$Estimates[3:4]
Y_c <- (Y - beta_hat[1]) / beta_hat[2]
fitted_model_c <- fit_model(x = X, r_x = r_vec_x, y = Y_c, r_y = r_vec_y, orders = c(1,1,2), order_select = FALSE)
library(doParallel, quietly = TRUE)
numCores <- detectCores()
registerDoParallel(numCores)
set.seed(123456789)
res_boot_c <- do_bootstrap(B = 10000, x = X, r_x = r_vec_x, y = Y_c, r_y = r_vec_y, orders = c(1,1,2))
```

Having calibrated the comparator measurements, refit the model with them, and rerun the bootstrap with them, we construct the bias and precision plots after calibration, by again calling `calculate_bias()` and `calculate_precision()`. Note that the plots below were not included in the main text of the paper.

```{r, fig.width=10, fig.height=3.75, fig.align='center'}
par(mfrow=c(1,3))
bias <- calculate_bias(theta = fitted_model_c$Estimates, orders = c(1,1,2), 
                       s = seq(from = min(fitted_model_c$BLA), to = max(fitted_model_c$BLA), length.out = 500),
                       interval_info = list(type = "standard", boot = res_boot_c, alpha = 0.05), 
                       plot_info = list(leg_location = "bottomleft", colour = "red", s_name = "tPSA (μg/L)"))
precision_x <- calculate_precision(theta = fitted_model_c$Estimates, orders = c(1,1,2), method = "R", 
                       s = seq(from = min(fitted_model_c$BLA), to = max(fitted_model_c$BLA), length.out = 500),
                       interval_info = list(type = "standard", boot = res_boot_c, alpha = 0.05), 
                       plot_info = list(leg_location = "topleft", colour = "blue", s_name = "tPSA (μg/L)", 
                                        method_name = "Roche", ylimits = c(0, 1.6)))
precision_y <- calculate_precision(theta = fitted_model_c$Estimates, orders = c(1,1,2), method = "C", 
                       s = seq(from = min(fitted_model_c$BLA), to = max(fitted_model_c$BLA), length.out = 500),
                       interval_info = list(type = "standard", boot = res_boot_c, alpha = 0.05), 
                       plot_info = list(leg_location = "topleft", colour = "purple", s_name = "tPSA (μg/L)", 
                                        method_name = "Siemens", ylimits = c(0, 1.6)))
```

We similarly constructed PoA plots after calibration by calling `calculate_poa()` with the calibrated model and bootstrap. The plots below reproduce Figure 4 in the paper.

```{r, fig.width=10, fig.height=5, fig.align='center'}
s <- seq(from = min(fitted_model_c$BLA), to = max(fitted_model_c$BLA), length.out = 500)
par(mfrow = c(1,2))
# Fixed delta
poa_fix <- calculate_poa(theta = fitted_model_c$Estimates, delta = 4, orders = c(1,1,2), s = s,
                       interval_info = list(type = "standard", boot = res_boot_c, alpha = 0.05), 
                       plot_info = list(leg_location = "bottomleft", colour = "orange", s_name = "tPSA (μg/L)"))
title(main = bquote("PoA Plot ("*delta*"=4)"))
# Relative delta
poa_rel <- calculate_poa(theta = fitted_model_c$Estimates, delta = "15%", orders = c(1,1,2), s = s,
                       interval_info = list(type = "standard", boot = res_boot_c, alpha = 0.05), 
                       plot_info = list(leg_location = "bottomright", colour = "orange", s_name = "tPSA (μg/L)"))
title(main = bquote("PoA Plot ("*delta*"=15%)"))
```